Phylogenetics and the human microbiome.

### 2013-06-05 sbreview.tex
....

% morgan et al extraction
% PyNAST

%For arXiv, uncomment this block and comment the next block.
\documentclass{amsart}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{url}
\usepackage[round]{natbib}
\newcommand{\forarxiv}[1]{#1}
\newcommand{\notforarxiv}[1]{}

@@

Terminology: survey versus metagenome.
Definition of a single sample.
Despite their importance, I will not describe experimental techniques such as gnotobiotic mice to stay within scope.

\cite{hugenholtz2002review}
{Exploring prokaryotic diversity in the genomic era}

\subsection{Marker genes}
History-- bacteria?

Our knowledge of the microbial world is due to a large part derived from the work of Carl Woese and colleagues who pioneered the use of marker genes as a way to distinguish between microbial lineages.
Their work, and the scientists who followed them, used especially the 16S ribosomal gene.
This gene was chosen because...
They developed the first "universal" 16S PCR primers.
The great accomplishment of this is learning that archaea, although morphologically similar to bacteria, form their own separate lineage.

@@

While Woese and colleagues used primitive sequencing methods to gain information, modern researchers have the luxury of high throughput sequencing.
This can be

Despite the high-throughput and low cost of modern sequencing, inherent challenges remain for the use of population census by marker gene sequencing.
Most fundamentally, various microbes have different DNA extraction efficiencies, even with hard core protocols, meaning that the representation of marker gene sequences is not representative of the actual communities \cite{morganEaInVitroSimulatedMetagenome10}.
Furthermore, primer bias.
Current sequencing technology is limited to a length that is shorter than most genes, which limits the resolution of the analyses.
Different regions may be good for different groups of organisms.

@@

% alternative loci
\cite{case2007rpob}

@@

Knowledge of the impact of microbial communities has been

@@

@@

Much of this work has been done using the 16s ribosomal rRNA gene.
Exceptions.

@@

\cite{mcnabb2004hsp65}
{{A}ssessment of partial sequencing of the 65-kilodalton heat shock protein gene (hsp65) for routine identification of {M}ycobacterium species isolated from clinical sources}

@@

RNA copy number
\cite{morganEaInVitroSimulatedMetagenome10}

@@

Trait evolution not considered.

@@

Clinical applications,
\cite{clarridge2004}
{{I}mpact of 16{S} r{RNA} gene sequence analysis for identification of bacteria on clinical microbiology and infectious diseases}
Understand cause and effect of diseases.
Take an ecololgical view on atopic dermititis.

For routine identification of a single microbe in culture, MALDI-TOF kicks ass and is cheap.

The field has experienced a frenetic rate of expansion over the past xxx years, and sometimes it can feel like researchers in this area have forgotten some of the basic principles of study design, such as power calculations, that are well recognized in other areas of medical research.
Although the human microbiome has led to a lot of "hype," the microbes inside of us and research on them is here to stay.
Thus we can look forward to the field of human microbiome analysis settling down to a comfortable and mature middle age as an interesting intersection between ecology and medicine.

>>>>

% PI-CRUST

\documentclass{amsart}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{url}
\usepackage[round]{natbib}
\newcommand{\forarxiv}[1]{#1}
\newcommand{\notforarxiv}[1]{}

@@

Terminology: survey versus metagenome.
Definition of a single sample.
There are a number of non-sequence-based experimental techniques that have been crucial for the development of the field, such as manipulation of gnotobiotic mice, but these are beyond the scope of this paper.

Many of these methods cross over with microbial ecology in general.
Exploring prokaryotic diversity in the genomic era \cite{hugenholtz2002review}.

\subsection{Microbial community estimation using marker genes}
History-- bacteria?

Our modern knowledge of the microbial world is due to a large part derived from the work of Carl Woese and colleagues who pioneered the use of marker genes as a way to distinguish between microbial lineages.
Their work, and the scientists who followed them, used especially the 16S ribosomal gene.
This gene was chosen because...
They developed the first "universal" 16S PCR primers.
The great accomplishment of this is learning that archaea, although morphologically similar to bacteria, form their own separate lineage.

@@

While Woese and colleagues used primitive sequencing methods to gain information, modern researchers have the luxury of high throughput sequencing.
This can be done with a high level of multiplexing, making an explicit trade-off between depth of sequencing for each specimen and the number of specimens able to be put on the sequencer at the same time.
This leads to the question of how many sequences are needed to characterize the microbial diversity of a given environment.
For simply separating two rather different samples, rather few samples are needed \cite{kuczynski2010microbial}.

Despite the high-throughput and low cost of modern sequencing, inherent challenges remain for the use of population census by marker gene sequencing.
Most fundamentally, various microbes have different DNA extraction efficiencies, even with hard core protocols, meaning that the representation of marker gene sequences is not representative of the actual communities \cite{morgan2010metagenomic}.
Furthermore, primer bias.
Current sequencing technology is limited to a length that is shorter than most genes, which limits the resolution of the analyses.
Different regions may be good for different groups of organisms.

@@

Some groups have reported advantages to using alternate single-copy genes as markers, (e.g. \cite{case2007rpob},\cite{mcnabb2004hsp65}), however nothing has really come close to 16S.

@@

As we will see, 16S is used frequently as a way of measuring diversity.
Those used to microbial genetics may think this surprising, because the genetic repertoire of microbes may not be so much vertically inherited.
However, \cite{zaneveld}
This approach has recently been taken to its logical conclusion by trying to infer.

@@

As described above ``metagenome" means that data is sheared randomly.
This type of data has a complication above that taken from 16S survey data, which is that the genetic region sequenced is unknown in addition to the organism it was taken from.

Metagenomic data is typically used to infer functional information directly.

One approach to using metagenomic data is simply to subset to genes that are found in a large proportion of micro-organisms.
This is the approach taken by the team of Jonathan Eisen, starting first with AMPHORA \cite{wu2008amphora} and subsequently with PhyloSift (in preparation).

In parallel, MLTreeMap.

Because of the diversity of gene repertoire in microbes, the gene sets may have only limited overlap with one another.
One solution is assembly and then phylogenetic analysis.
metAMOS pipeline \cite{treangen2011metamos}.

A fascinating approach is that of \citet{segata2012metagenomic}, who start by compiling a database of clade-specific genes \citep{segata2011metagenomic}.
Then they classify the origin of a given read as being the only clade that has this gene.
They show that this has good specificity, however, sensitivity is limited to genomes that have been sequenced.

@@

@@

@@

RNA copy number
\cite{morgan2010metagenomic}

@@

@@

Clinical applications,
\cite{clarridge2004}
{{I}mpact of 16{S} r{RNA} gene sequence analysis for identification of bacteria on clinical microbiology and infectious diseases}
For routine identification of a single microbe in culture, MALDI-TOF kicks ass and is cheap.
Diagnostics involving mixed communities.
Bioinformatics will get hardened into fixed workflows.

Understand cause and effect of diseases.
Take an ecololgical view on atopic dermititis.

Microbial outbreak tracking.
Very common in RNA viruses.
Already using similar methods by applying whole-genome sequencing.

The field has experienced a frenetic rate of expansion over the past xxx years, and sometimes it can feel like researchers in this area have forgotten some of the basic principles of study design, such as power calculations, that are well recognized in other areas of medical research.
Although the human microbiome has led to a lot of "hype," our microbes are here to stay and so is research on them.
Thus we can look forward to the field of human microbiome analysis settling down to a comfortable and mature middle age as an interesting intersection between ecology and medicine.
